Manifold Clustering for Latent Structure Block Models

### 2022-03-05 draft/draft.Rmd
fixed a few bugs

```{r, cache = TRUE}
maxit <- 50
animate <- FALSE
animation.dir <- '~/dev/manifold-block-models/draft'
random.starts <- 2
eps <- 1e-3
out <- lapply(seq(random.starts), function(i) {
  set.seed(i)
  intercept <- (i / random.starts > .5 + eps)
  manifold.clustering(Xhat, 2, 
                      parallel = TRUE, 
                      intercept = intercept, 
                      maxit = maxit,
                      eps = 1e-3,
                      verbose = FALSE,
                      animate = animate,
                      animation.dir = animation.dir,
                      animation.title = i)
})
names(out) <- paste('random', seq(random.starts))
names(out) <- sapply(seq(random.starts), function(i) {
  if (i / random.starts < .5 + eps) {
    paste('random', i, '\n(w/o intercept)')
  } else {
    paste('random', i, '\n(w/ intercept)')
  }
})

@@

  t. <- seq(0, 1, .01)
  Y1 <- bezier.curve(t., out[[i]]$p[[1]])
  Y2 <- bezier.curve(t., out[[i]]$p[[2]])

@@

```{r cache = TRUE}
set.seed(123)
z.spectral <- spectral.clustering(A)
macaque.out <- manifold.clustering(Xhat, 2, 
                                   initialization = z.spectral, 
                                   intercept = TRUE, 
                                   parallel = TRUE, 
                                   verbose = FALSE, 
                                   animate = animate, 
                                   curve.init = 'random',
                                   animation.title = 'macaque')
```

```{r}
t.1 <- seq(min(macaque.out$t[[1]]), max(macaque.out$t[[1]]), .01)
t.2 <- seq(min(macaque.out$t[[2]]), max(macaque.out$t[[2]]), .01)
Y1 <- bezier.curve(t.1, macaque.out$p[[1]])
Y2 <- bezier.curve(t.2, macaque.out$p[[2]])

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], 
                 colour = factor(macaque.out$z)),
             size = .5) + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  coord_fixed() +
  theme(legend.position = 'none') +
  geom_path(aes(x = Y1[, 1], y = Y1[, 2])) +
  geom_path(aes(x = Y2[, 1], y = Y2[, 2])) + 
  xlim(min(Xhat[, 1]), max(Xhat[, 1])) + ylim(min(Xhat[, 2]), max(Xhat[, 2]))
```

@@

plot(data.matrix, col = z * 2, asp = 1, xlab = NA, ylab = NA,
     main = 'Latent configuration')
```

```{r}
P <- data.matrix %*% diag(c(1, 1)) %*% t(data.matrix)
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
plot(Xhat, col = z * 2, asp = 1, main = 'ASE')
```

>>>>

```{r, cache = TRUE}
maxit <- 50
animate <- FALSE
animation.dir <- '~/dev/manifold-block-models/draft'
random.starts <- 2
eps <- 1e-3
out <- lapply(seq(random.starts), function(i) {
  intercept <- (i / random.starts > .5 + eps)
  set.seed(i)
  manifold.clustering(Xhat, 2, 
                      parallel = TRUE, 
                      intercept = intercept, 
                      maxit = maxit,
                      eps = 1e-3,
                      verbose = FALSE,
                      animate = animate,
                      animation.dir = animation.dir,
                      animation.title = i)
})
names(out) <- paste('random', seq(random.starts))
names(out) <- sapply(seq(random.starts), function(i) {
  if (i / random.starts < .5 + eps) {
    paste('random', i, '\n(w/o intercept)')
  } else {
    paste('random', i, '\n(w/ intercept)')
  }
})

@@

  t.1 <- seq(min(out[[i]]$t[[1]]), max(out[[i]]$t[[1]]), .01)
  t.2 <- seq(min(out[[i]]$t[[2]]), max(out[[i]]$t[[2]]), .01)
  Y1 <- bezier.curve(t.1, out[[i]]$p[[1]])
  Y2 <- bezier.curve(t.2, out[[i]]$p[[2]])

@@

```{r cache = TRUE}
# set.seed(123)
z.spectral <- spectral.clustering(A)
macaque.out <- manifold.clustering(Xhat, 2, 
                                   # initialization = 'random',
                                   # initialization = z.spectral, 
                                   initialization = z,
                                   intercept = FALSE, 
                                   parallel = TRUE, 
                                   verbose = FALSE, 
                                   animate = animate, 
                                   curve.init = 'x',
                                   animation.title = 'macaque')
```

```{r}
t.1 <- seq(min(c(0, macaque.out$t[[1]])), max(macaque.out$t[[1]]), .01)
t.2 <- seq(min(c(0, macaque.out$t[[2]])), max(macaque.out$t[[2]]), .01)
Y1 <- bezier.curve(t.1, macaque.out$p[[1]])
Y2 <- bezier.curve(t.2, macaque.out$p[[2]])

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], 
                 colour = factor(macaque.out$z)),
             size = .5) + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  coord_fixed() +
  theme(legend.position = 'none') +
  geom_path(aes(x = Y1[, 1], y = Y1[, 2])) +
  geom_path(aes(x = Y2[, 1], y = Y2[, 2]))
```

@@

plot(data.matrix, col = z * 2, asp = 1, xlab = NA, ylab = NA)
```

```{r}
P <- data.matrix %*% diag(c(1, 1)) %*% t(data.matrix)
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
# plot(Xhat, col = z * 2, asp = 1)
```
